---
title: "Interesting Plots"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("tidyup.R")
#now health stores the data. SPlit it up so we have a exploratory set, a query set, and a confirmation set.

set.seed(927)
sample_inds<-seq(1,nrow(health))
train_set<-sample(sample_inds,floor(.6*nrow(health)))

#remove the indices we already sampled
rem_sample_inds<-sample_inds[!(sample_inds %in% train_set)]
#build the confirmation and query set from the remaining indices
confirmation_set<-sample(rem_sample_inds,.5*length(rem_sample_inds))
query_set<-rem_sample_inds[!(rem_sample_inds %in% confirmation_set)]

health_full<-health
#break up the table. Store the training set in health so the rest of the code works.
health_query<-health[query_set,]
health_confirm<-health[confirmation_set,]
health<-health[train_set,]
```


```{r, echo=FALSE}
plot_Missing <- function(data_in, title = NULL){
  temp_df <- as.data.frame(ifelse(is.na(data_in), 0, 1))
  temp_df <- temp_df[,order(colSums(temp_df))]
  data_temp <- expand.grid(list(x = 1:nrow(temp_df), y = colnames(temp_df)))
  data_temp$m <- as.vector(as.matrix(temp_df))
  data_temp <- data.frame(x = unlist(data_temp$x), y = unlist(data_temp$y), m = unlist(data_temp$m))
  ggplot(data_temp) + geom_tile(aes(x=x, y=y, fill=factor(m))) + scale_fill_manual(values=c("white", "black"), name="Missing\n(0=Yes, 1=No)") + theme_light() + ylab("") + xlab("") + ggtitle(title)
}

plot_Missing(health)
```

```{r, echo=FALSE}
health %>%
  filter(BIO_SEX!="Refused") %>%
  ggplot() +
  geom_histogram(aes(CGPATOT,..density..),binwidth=.25) +
  facet_wrap(~BIO_SEX)
```
```{r, echo=TRUE}
dh_visits <- health %>%
  filter(!(DENTXMW1 %in% c("Refused", "Don't know", "Not applicable")),
         !(PHYSXMW1 %in% c("Refused", "Don't know", "Not applicable")))

dh_visits %>%
  group_by(DENTXMW1,PHYSXMW1) %>%
  summarise(n=n())

dh_visits %>%
  ggplot() +
  geom_histogram(aes(CGPATOT,..density..),binwidth=.25) +
  facet_grid(DENTXMW1 ~ PHYSXMW1)
```


Plot that breaks down race versus health

```{r, echo=FALSE}
health %>%
  filter(GENHTHW1 %in% c("Excellent","Very Good","Good","Fair","Poor") & !is.na(FAMRACE)) %>%
  ggplot() + 
  geom_bar(aes(x=GENHTHW1,y=..prop..,group=1)) +
  facet_wrap(~FAMRACE)
```

```{r, echo=FALSE}

health %>%
  ggplot() +
  geom_bar(aes(x=PARENTS))
```

```{r, echo=FALSE}
health %>%
  ggplot() +
  geom_density(aes(CGPATOT)) +
  facet_wrap(~PARENTS)
```

```{r, echo=FALSE}
health %>%
  ggplot() +
  geom_bar(aes(CGPASTEM)) +
  facet_wrap(~PARENTS)
```

A Two-Way ANOVA of parents effect on GPA
```{r, echo=TRUE}
modvar<-health %>%
  select(CGPASTEM,HASMOM,HASDAD) %>%
  filter(!is.na(CGPASTEM)) %>%
  group_by(HASMOM,HASDAD)

pmod<-lm(CGPASTEM ~ HASMOM + HASDAD, data=modvar)
anova(pmod)
```


```{r echo=TRUE}
health %>% 
  group_by(PARENTS) %>%
  summarise(GPAVG=mean(CGPATOT,na.rm=TRUE))
```

```{r echo=TRUE}
summary(lm(CGPATOT ~ FAMRACE, data=select(health,-CGPASTEM,-AID,-CAGE)))
```



```{r echo=TRUE}
summary(lm(CGPATOT ~ CGPASTEM, data=health))
```

```{r echo=FALSE}
health %>% group_by(PHYSXMW1,DENTXMW1) %>% summarise(n=n()) %>% ggplot(aes(x=PHYSXMW1,DENTXMW1)) + geom_tile(aes(fill=n)) + geom_label(aes(label=n))
```

This data does not seem like a representative sample of the actual high school population. It seems that, in particular, African Americans are way underrepresented. Enrollment of African Americans in school was 15% in 2008 (source: https://www.census.gov/prod/2011pubs/p20-564.pdf). Even though enrollment proportions are non-stationary, there shouldn't be a 25x increase from 1994/5 to 2008 of the enrollment of African American students. In addition, Native American students seem over represented, but I don't have a source to back this up.
```{r echo=TRUE}
health %>%
  ggplot() +
  geom_bar(aes(FAMRACE)) +
  coord_flip()

```

```{r, echo=TRUE}
health_full %>%
  group_by(FAMRACE) %>%
  summarise(p=n()/nrow(health_full))
```
Analyzing the number of NAs in GPA based off race
```{r echo=TRUE}
health %>%
  group_by(FAMRACE) %>%
  summarise(sum(is.na(CGPATOT))/n())
```

```{r echo=TRUE}
health %>%
  group_by(FAMRACE) %>%
  summarise(mean(CGPATOT,na.rm=TRUE))
```

```{r echo=TRUE}
health %>%
  group_by(DADEDW1) %>%
  summarise(mean(CGPATOT,na.rm=TRUE))
```

```{r echo=TRUE}
health %>%
  group_by(DADWKW1) %>%
  summarise(mean(CGPATOT,na.rm=TRUE), n=n())
```

```{r echo=TRUE}
health %>%
  group_by(DADWKW1) %>%
  summarise(mean(CGPASTEM,na.rm=TRUE), n=n())
```