---
title: "Extracting Indicators of Adolescent Depression using Multivariate Regression"
author: |
  | Team Infinite Limits: 
  | Jeffrey Limbacher, Robert Cholvin, and Evan Habbershaw
date: "November 3, 2017"
output: beamer_presentation
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rpart)
library(DMwR2)
library(randomForest)
library(rpart.plot)
library(leaps)
source("label_data.R")

#from http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Teen Depression

- A study published in *Pediatrics* showed that from 2005 to 2014, the number of teenage major depressive episodes increased from 8.7% in 2005 to 11.5 % in 2014. \footnote{Kataoka SH, Zhang L, Wells KB, "Unmet need for mental health care among U.S. children: variation by ethnicity and insurance status"}

- NPR reports that mental health problems can manifest as  chronic absence, low achievement, disruptive behavior and dropping out. \footnote{Anderson, "Mental Health In Schools: A Hidden Crisis Affecting Millions Of Students"}

- The NPR story follows a student who was diagnosed with depression and bulimia.

- "I felt like nobody wanted to help me".

- What are indicators of depression?

## Analyzing Depression within the ADD health set


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
p1 <- health %>%
  ggplot(aes(x=DEPREW1)) +
  geom_histogram(bins = 25) +
  geom_point(aes(x=DEPREW1,y=0)) +
  labs(x="Depression Wave") +
  ggtitle("Histogram of Depression")

health2 <- health %>%
  filter(!is.na(DEPREW1) & DEPREW1 > 0) %>%
  mutate(DEPREW1=log(DEPREW1))

p2 <- health2 %>%
  ggplot() +
  geom_qq(aes(sample=DEPREW1)) +
  ggtitle("QQ-Normal Plot vs log(Depression)")

multiplot(p1,p2, cols = 1)
```

## Multivariable Regression: Residual analysis

- We performed a multivariable regression of depression.

```{r}
predictors <- unique(c("INTELLW1" , "MFUTEXPW1" , "DFUTEXPW1", "CGPATOTW1",
                  "NEGLFEXPW1", "POSLFEXPW1" , "SELFESTW1" , "MENTHFREW1" , "WLUNHLTHW1" ,
                  "BODYIMGW1" , "MOMEDW1" , "MOMWKW1" , "DADWKW1" , 
                  "IMPRELW1" , "SOCSPPTW1" , "H1GH48", "ADOPTW1",
                  "POSFMENVW1", "BIO_SEX", "SLEEP","SCLBELNG2W1","SCHLTRBLW1","CPRBEHLMW1",
                  "SCHEXPEL","PHYHFREQW1","CRPWBW1","CSCLBELNGW1",
                  "DLYACT1W1","DLYACT2W1","DLYACT3W1","DLYACT4W1","DLYACT5W1","DLYACT6W1","DLYACT7W1",
                  "CLUBACADW1","CLUBMUARTW1","CLUBSPORTW1"))


health_rmna <- na.omit(health[,c("DEPREW1",predictors)])
health_rmna <-droplevels.data.frame(health_rmna)
health_rmna <- health_rmna %>%
  mutate_at(vars(INTELLW1,BODYIMGW1,IMPRELW1,H1GH48,starts_with("DLYACT"),starts_with("CLUB")), as.numeric)
health_rmna <- health_rmna %>%
  filter(DEPREW1 > 0) %>%
  mutate(DEPREW1 = log(DEPREW1))


health_rmna <- health_rmna[c(-337,-382,-172,-916),]

form<-as.formula(paste("DEPREW1 ~ ", paste(c(predictors,"BODYIMGW1*BIO_SEX"),collapse="+")))

depmod <- lm(as.formula(paste("DEPREW1 ~ ", paste(c(predictors,"BODYIMGW1*BIO_SEX"),collapse="+"))),
                data=health_rmna, na.action=na.omit)

depmod_both<-step(depmod,direction="both",trace=0)

par(mfcol=c(1,2))

plot(depmod_both, which=c(1,2))
```

## Significant Variables

- We used R's _lm_ and _step_ function to find the most important variables.
- $R^2=.46$

```{r echo=FALSE, results="asis"}
library(knitr)
library(pander)
knitr::opts_chunk$set(size = "tiny")

dep_sum <- summary(depmod_both)$coefficients

coeffs <- as.tibble(cbind(Variable=names(depmod_both$coefficients),dep_sum)) %>%
  mutate_at(vars(`Estimate`:`Pr(>|t|)`),as.double) %>%
  arrange(`Pr(>|t|)`) %>% 
  mutate_at(vars(`Estimate`:`Pr(>|t|)`), function(x) round(x,digits=3))  %>%
  filter(`Pr(>|t|)` < .01) %>%
  filter(Variable!="(Intercept)") %>%
  select(-`Std. Error`)

coeffs[1,"Variable"]<-"Mental Health Concerns"
coeffs[2,"Variable"]<-"Self Esteem"
coeffs[3,"Variable"]<-"Trouble in School"
coeffs[4,"Variable"]<-"School absences"
coeffs[5,"Variable"]<-"Enough Sleep"
coeffs[6,"Variable"]<-"Positive Family Environment"
coeffs[7,"Variable"]<-"Total GPA"
coeffs[8,"Variable"]<-"Religion Importance"
coeffs[9,"Variable"]<-"Hobbies"
coeffs[10,"Variable"]<-"Health Issues"

kable(coeffs)
```


## Conclusions

- Our results provide anyone who works with adolescents indicators to look out for.
- Limitations:
    - Our model had an $R^2$ of .46 which indicates that there are more unexplained variables.
    - We didn't incorporate interaction terms.
    - There were many observations with missing values which were excluded (6504 $\rightarrow$ 1621).
- Our work can be found at github.com/jeffreylimbacher/csun-data-jam